---
title: "Peer Assessment 1"
author: "Sai"
date: "Saturday, May 16, 2015"
output: html_document
---

## Part 1: Loading and preprocessing the data

```{r,echo=TRUE,cache=TRUE}
raw.data<-read.csv(file.choose(),header=T)
head(raw.data)
summary(raw.data) 
```

## Part 2: What is mean total number of steps taken per day?

```{r,echo=TRUE,cache=TRUE}
options(scipen=1,digits=2)
data1<-raw.data[!is.na(raw.data$steps),]
data1$date<-factor(data1$date)

library(plyr)
sum.day<-ddply(data1,.(date),summarise,total=sum(steps))

library(ggplot2)
g<-ggplot(sum.day,aes(sum.day$total))
g<-g+geom_histogram(col="red",fill="green",alpha=0.4,binwidth=1200)
g<-g+xlab("Total Steps Each Day")+ylab("Frequency")
g<-g+ggtitle("Histogram of the Total Number of Steps Taken Each Day")
g

ave<-mean(sum.day$total)
med<-median(sum.day$total)
```
The mean of total number of steps taken per day is `r ave`.

The median of total number of steps taken per day is `r med`.

## Part 3: What is the average daily activity pattern?

```{r,echo=TRUE,cache=TRUE}
options(scipen=1,digits=2)
ave.int<-ddply(data1,.(interval),summarise,average=mean(steps))
max.ave<-max(ave.int$average)
max.int<-ave.int[ave.int$average==max.ave,1]
max.int.adj<-max.int-5
g<-ggplot(ave.int,aes(x=interval,y=average))+geom_line()
g<-g+geom_segment(aes(x=max.int,y=0,xend=max.int,yend=max.ave),col="red")
g
```

The maximum number of steps is `r max.ave`.

It occurse between `r max.int.adj` minute and `r max.int` minute.

## Part 4: Imputing missing values
```{r,echo=TRUE,cahce=TRUE}
n.na<-sum(is.na(raw.data$steps))
```
The total number of missing values in the dataset is `r n.na`.

Fill in the missing values with the mean of taht 5-minute interval.
```{r,echo=TRUE,cache=TRUE}
N<-nrow(raw.data)
attach(raw.data)
tstep<-numeric(N)
for (i in 1:N){
  if(is.na(steps[i])) tstep[i]=ave.int[ave.int$interval==interval[i],2]
	else tstep[i]=steps[i]
}
detach(raw.data)
data2<-raw.data
data2$steps<-tstep

sum.day2<-ddply(data2,.(date),summarise,total=sum(steps))
g<-ggplot(sum.day2,aes(sum.day2$total))
g<-g+geom_histogram(col="red",fill="green",alpha=0.4,binwidth=1200)
g<-g+xlab("Total Steps Each Day")+ylab("Frequency")
g<-g+ggtitle("Missing Values Adjusted Histogram")
g

ave2<-mean(sum.day2$total)
med2<-median(sum.day2$total)
```

After filling in the missing values:

The mean total number of steps taken per day is `r ave2`,
which is the same as the number in part 1.

The median total number of steps taken per day is `r med2`,
which is slightly higher than the one in part 1.

## Part 5: Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE,cache=TRUE}
tdate<-weekdays(as.Date(data2$date))
dayfact<-character(N)
for (i in 1:N){
  if(tdate[i]=="Saturday"|tdate[i]=="Sunday") dayfact[i]="weekend"
	else dayfact[i]="weekday"
}
dayfact<-factor(dayfact)

data3<-data.frame(data2,dayfact)
aveday<-ddply(data3[dayfact=="weekday",],.(interval),
              summarise,average=mean(steps))
aveday$dayfact<-"weekday"
aveend<-ddply(data3[dayfact=="weekend",],.(interval),
              summarise,average=mean(steps))
aveend$dayfact<-"weekend"

data4<-rbind(aveday,aveend)

g<-ggplot(data4,aes(x=interval,y=average))
g<-g+facet_grid(dayfact~.)+geom_line(aes(colour=dayfact))
g
```

It seems the activity patterns between weekdays and weekends are a little bit different.